---
title: "assignment_2"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.height = 3, 
                      fig.width  = 6,
                      fig.align  = "center")
ggplot2::theme_set(ggplot2::theme_bw())
```

```{r}
#| message: false
library(tidyverse)

library(lubridate)
```


# Question 1:

Write a function `pos_na()` that takes two vectors of equal length and returns the positions where **both** vectors contain `NA`.  
If the vectors do **not** have the same length, the function should return the message:

> `"The vectors are not the same length."`

Use the following pairs of vectors to test your function:

1. `c(1, NA, 2)` and `c(NA, NA)`
2. `c(NA, NA, 2)` and `c(NA, NA, 2)`
3. `c(NA, 5, NA)` and `c(NA, NA, NA)`
4. `c(NA, NA, NA, 2, NA, 4, NA)` and `c(NA, NA, 2, 4, 4, NA, NA)`
5. `c(1, NA, NA, 2, NA, 4, NA)` and `c(NA, -5, 2, 4, 4, NA, 11)`


```{r}
pos_na <- function(x, y) {
  if (length(x) != length(y)) {
    return("The vectors are not the same length.")
  }
  which(is.na(x) & is.na(y))
}
```

```{r}
pos_na(c(1, NA, 2), c(NA, NA))
```
```{r}
pos_na(c(NA, NA, 2), c(NA, NA, 2))
```

```{r}
pos_na(c(NA, 5, NA), c(NA, NA, NA))
```

```{r}
pos_na(
  c(NA, NA, NA, 2, NA, 4, NA),
  c(NA, NA, 2, 4, 4, NA, NA)
)
```


```{r}
pos_na(
  c(1, NA, NA, 2, NA, 4, NA),
  c(NA, -5, 2, 4, 4, NA, 11)
)
```

```{r}
pos_na2(c(NA, NA, 2), c(NA, NA, 2))
```

```{r}
pos_na2(c(NA, 5, NA), c(NA, NA, NA))
```

```{r}
pos_na(
  c(NA, NA, NA, 2, NA, 4, NA),
  c(NA, NA, 2, 4, 4, NA, NA)
)
```

```{r}
pos_na(
  c(1, NA, NA, 2, NA, 4, NA),
  c(NA, -5, 2, 4, 4, NA, 11)
)
```


### (1ii) Without using an `if` statement

Write a second version of `pos_na()` that performs the same task **without** using an `if` statement.

-   You may use any other valid R mechanism (e.g., `stop()`, `stopifnot()`, or logical short‑circuiting) to handle the case where the vectors differ in length.

```{r}
pos_na2 <- function(x, y) {
  length(x) == length(y) ||
    stop("The vectors are not the same length.")
   which(is.na(x) & is.na(y))
}

```

```{r}
#pos_na2(c(1, NA, 2), c(NA, NA))
#Error: The vectors are not the same length
```


Here’s a polished, professional way to label those two parts of the assignment. It reads cleanly and sets expectations without adding extra wording.

---

### (1i) Using an `if` statement

Write a version of `pos_na()` that checks whether the two input vectors have the same length using an `if` statement.  

- If they do, return the positions where both vectors contain `NA`.  

- If they do not, return the message:

> `"The vectors are not the same length."`



### (1ii) Without using an `if` statement

Write a second version of `pos_na()` that performs the same task **without** using an `if` statement.  

- You may use any other valid R mechanism (e.g., `stop()`, `stopifnot()`, or logical short‑circuiting) to handle the case where the vectors differ in length.

``
  
# Question 2
 Load the `wmata_ridership` data frame into R from <https://dcgerard.github.io/stat_412_612/data/wmata_ridership.csv>. For each month, calculate the proportion of rides made on a given day of the month. Then make box plots of the proport
```{r}
library(tidyverse)
library(lubridate)

ridership <- read_csv(
  "https://dcgerard.github.io/stat_412_612/data/wmata_ridership.csv",
  col_types = cols(.default = "c")
)

#clean
ridership <- ridership |>
  mutate(
    Date  = as.Date(Date),
    Total = as.numeric(gsub(",", "", Total)), 
    year  = year(Date),
    month = month(Date),
    dow   = wday(Date, label = TRUE, week_start = 1))

# remove 2004 and 2005
ridership_filtered <- ridership |>
  filter(!year %in% c(2004, 2005))

# monthly totals
monthly_totals <- ridership_filtered |>
  group_by(year, month) |>
  summarize(monthly_total = sum(Total, na.rm = TRUE), .groups = "drop")

# daily proportions
ridership_prop <- ridership_filtered |>
  left_join(monthly_totals, by = c("year", "month")) %>%
  mutate(prop = Total / monthly_total)

ggplot(ridership_prop, aes(x = dow, y = prop)) +
  geom_boxplot(fill = "hotpink", alpha = 0.7) +
  labs(
    title = "Ridership Proportions by Day of the Week",
    x     = "Day of Week",
    y     = "Daily proportion of monthly ridership"
  ) +
  theme_minimal()

```
# Question 3:
Write R code that extracts the elements "Bears", "Dolphins", and "Bengals" from the vector `V` shown below, and display the result.
```{r}
V <- c("Bears", "Lions", "Dolphins", "Eagles", "Bengals")
```


## (3i) Subsetting by index
```{r}
V[c(1, 3, 5)]
```
## (3ii) Subsetting using a `for` loop
```{r}
indices <- c(1, 3, 5)
result <- c()

for (i in indices) {
  result <- c(result, V[i])
}

result
```

## (3iii)Returning all values in a single vector (one line of code)
```{r}
V[c(1, 3, 5)]
```
### (a) Using positive indices
```{r}
V[c(1, 3, 5)]
```
### (b) Using negative indices
```{r}
V[-c(2, 4)]
```
### (c) Using logical indexing
```{r}
V[c(TRUE, FALSE, TRUE, FALSE, TRUE)]
```
### (d) Attempting to subset by names
```{r}
names(V) <- c("A", "B", "C", "D", "E")
V[c("A", "C", "E")]
```
# Question 4:

You are given a character vector containing messy employee records. Each entry includes:

- an employee name (string),
- a hire date in inconsistent formats,
- and a department code that should be treated as a factor.

```{r}
records <- c(
  "Steve McQueen | 2020-01-15 | HR",
  "B. Smith | 15/02/2021 | FIN",
  "Carlos M | March 3, 2019 | IT",
  "D. Lee | 2018/07/30 | HR",
  "Alain Delon | 04-12-2020 | MKT"
)
records
```

1. **Parse each record** into three separate variables:  
   - `name`  
   - `hire_date`  
   - `dept
```{r}   `
split_records <- strsplit(records, " \\| ")
parsed_records <- tibble(record = records) |>
  separate(record,
           into = c("name", "hire_date", "dept"),
           sep = " \\| ") |>
  mutate( name = str_trim(name),
    hire_date = str_trim(hire_date),
    dept = str_trim(dept))

parsed_records
```

2. **Convert `hire_date` into a proper `Date` object**, correctly handling all of the different date formats.
```{r}
library(lubridate)
parsed_records <- parsed_records |>
  mutate(
    hire_date = coalesce(
      ymd(hire_date, quiet = TRUE), 
      dmy(hire_date, quiet = TRUE),       
      mdy(hire_date, quiet = TRUE),     
      ymd(str_replace_all(hire_date, "/", "-"), quiet = TRUE)))

parsed_records

```

3. **Convert `dept` into an ordered factor** with the levels:  
   `HR`, `FIN`, `IT`, `MKT`.
```{r}
parsed_records <- parsed_records |>
  mutate(dept = factor( dept,
      levels = c("HR", "FIN", "IT", "MKT"),
      ordered = TRUE))

parsed_records
```

4. **Construct a clean data frame** containing the parsed and converted variables.

```{r}
clean_employees <- parsed_records |>
  select(name, hire_date, dept)

clean_employees
```

5. **Add a new variable** called `years_worked` that reports the number of full years each employee has worked as of today.


```{r}
clean_employees <- clean_employees |>
  mutate(years_worked = interval(hire_date, today()) %/% years(1))

```

6. **Return the final cleaned data frame.**

```{r}
clean_employees
```
